---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# GPsim: An R Package for Simulating Gaussian Processes

The R package `GPsim` is a simulation framework for generating realizations of both stationary and nonstationary Gaussian Processes (GPs), which allows users to make use of a wide range of simulation designs including arbitrary domains and user-defined covariance models.


## Installation

```{install}
# install.packages("devtools")
devtools::install_github("myeongjong/GPsim")

# If you need vignettes ...
# devtools::install_github("myeongjong/GPsim", build_vignettes = T)
```


## Example

```{r example, message=FALSE}
library(GPsim)

set.seed(2019)

covparms <- c(1, 0.1, 10) # (variance, range, degree of anisotropy)

realization <- simulate_gp(locs = "random", n = 15^2, p = 2, meanmodel = function(locs, meanparms) rep(0, nrow(locs)), meanparms = NULL, covmodel = cov_expo_aniso, covparms = covparms, pivot = FALSE, correction = NULL, tol = .Machine$double.eps, seed = 2020)

fields::quilt.plot(realization$locs[, 1], realization$locs[, 2], realization$y, main = "A realization of the standard GP", xlab = "loc1", ylab = "loc2")
```


## License

This package is free and open source software, licensed under GPL (>=2).
